Red-Giant Hot Jupiters: Brilliant Radio Emitter?

### 2015-04-24 ms.tex
figure replaced

%\subsection{}
The peak of the auroral radio emission is around the cyclotron frequency, of the planetary magnetic field $f_{\rm cyc}$: 
%In our assumption, the frequency band width is assumed to be scaled with the cyclotron frequency, $f_{\rm cyc}$: 
%%%
\begin{eqnarray}
f_{\rm cyc} &=& \frac{eB}{2\pi m_e c} \approx 28 {\rm~MHz} \left( \frac{B}{10 \rm~G} \right) \, ,
\label{eq:cyc} \\
B &=& \frac{\mu_0}{2\pi}\frac{\mathcal{M}}{R_p^3} \\
&\sim & 9.1{\rm~G} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right) \left( \frac{R_p}{R_{p, J}} \right)^{-3}. 
\end{eqnarray}
%%%
%where $\mathcal{M}_J = 1.56 \times 10^{27} \mbox{A m}^2$. 
For magnetic field strengths in the vicinity of $\sim$25-50~Gauss (G),
the cyclotron frequency will be in the range $\sim$75-150~MHz.

Note that the emission is observable from ground only when the maximum frequency is larger than the plasma frequency of the Earth's ionosphere, $f_{\rm c1}\sim 10$~MHz {\it and} the maximum plasma frequency along the line of sight, $f_{\rm c2}$. 
At the favorable condition where the planet is in the near side of the spherical plane, the maximum plasma frequency corresponds to that in the vicinity of the planet, $f_{\rm plasma}$. Therefore, 
%%%
\begin{eqnarray}
f_{\rm c2} &=& f_{\rm plasma} = \frac{1}{2\pi} \sqrt{\frac{ne^2}{\epsilon _0 m_e}} \\
&\approx & 4{\rm~MHz} \left( \frac{\dot M}{10^6 \dot M_{\odot}}\right)^{1/2} \left(\frac{v}{10^{-1} v_{\odot}}  \right)^{-1/2}
\end{eqnarray}
%%%

@@

Unlike usual hot jupiters, RGHJs are not subject to tidal lock, as the gravitational effects of their host star does not change even if the star evolves into red giants. Without no physical insights of the typical spin period, we simply assume that of Jupiter: $\omega = 9.925$ [hr].

In order to evaluate $\rho _c $ and $r_c$, we need a model of internal structure of gaseous planets. 
First, we assume that the planetary radius is constant at $R_p = R_{p,J}$, as the numerical calculations show that the radii of gaseous planets over the range of $0.1 R_{p, J} < M_p < 10M_{p, J}$ (with core mass less than 10\%) are converged to $0.8 R_{p, J} < R_p < 1.2R_{p, J}$ in 1 Gyr \citep{fortney2007}. 
For the density profile, we assume a polytrope gas sphere with index $n=1$, which results in:
%%%%%%%%%% 
\begin{equation}
\rho (r) = \left( \frac{\pi M_p}{4 R_p^3} \right) \frac{\sin \left( \pi \frac{r}{R_p} \right)}{\left( \pi \frac{r}{R_p} \right)}. \label{eq:rho_r}
\end{equation}
%%%%%%%%%%
We determine the core radius $r_c$ by assuming that the hydrogen becomes metallic when $\rho (r)$ exceeds the critical density $\rho_c=700\,\mbox{kg/m}^3$. The density of the metallic core, $\rho _c$ is obtained by averaging the density in the core. 
We assume that the conductivity $\sigma $ is the same as Jupiter \memoYF{?}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[htdp]
\caption{Canonical Parameters of 2 types of hot jupiters.}
\begin{center}
\begin{tabular}{c|ccc} \hline \hline
& mass & orbital distance &  spin period \\ 
& $M_p$ & $a$ & $\omega $ \\ \hline
canonical HJ & $M_J$ & 0.05 AU & 4 days \\
RGHJ & $M_J$ & 5 AU &  0.4 days \\ \hline
\end{tabular}
\end{center}
\label{tab:comp_HJ}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

On the assumption that the radio emission is proportional to the input kinetic energy (equation (\ref{eq:Pinp_kin})), the scaling of the radio emission is expanded as follows: 
%%%
\begin{eqnarray}
P_{\rm k,\,inp} &=& nv^3 r_s^2 = n v^3 \left( \frac{\mu_0^2 \mathcal{M}^2}{8\pi^3 m_{\rm p}nv^2} \right)^{1/3}  \\
&=& P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{2/3} \\
&& \times \left( \frac{\dot M}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \left( \frac{a}{5.2~\mbox{AU}} \right)^{-4/3} \label{eq:Pkinp}
\end{eqnarray}
%%%

It is illustrative to compare radio emission power of canonical hot jupiters and that of red giant hot jupiters. We adopt canonical values for planetary mass, orbital distance, and orbital period of each type as shown in Table \ref{tab:comp_HJ}. Then, equation (\ref{eq:Pkinp}) can be re-normalized as 
%%%
\begin{eqnarray}
P_{\rm k,\,inp} 
&\approx & 220 ~P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{2/3} \\
&& \times \left( \frac{\dot M}{10^6 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \left( \frac{a}{5.2~\mbox{AU}} \right)^{-4/3} \\
&\approx & 2.2 \times 10^4 ~P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{2/3} \\
&& \times \left( \frac{\dot M}{10^9 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \left( \frac{a}{5.2~\mbox{AU}} \right)^{-4/3} \\
&\approx & 100 ~P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{0.1\mathcal{M}_J} \right)^{2/3} \\
&& \times \left( \frac{\dot M}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \left( \frac{a}{0.05~\mbox{AU}} \right)^{-4/3} \\
\end{eqnarray}
%%%
\memoYF{what is the typical wind velocity of AGB stars?}
Thus, the jupiters to the AGB stars are intrinsically 200 times as bright as canonical hot Jupiters. In other words, we could search for radio signal from these stars in $(\sqrt{200})^3 \approx 3000$ times as large volume as the canonical hot jupiters.

We now turn to the spectral radio flux observed at the Earth. 
%%%
\begin{eqnarray}
F_{\nu} &=& \frac{P_{\rm radio}}{\Omega d^2 f_{\rm cyclotron}} \\
%&=& P_{\rm radio,\,J} \\
&\approx & 5\times 10^{-8} \mbox{Jy} \left( \frac{d}{100\mbox{~pc}} \right)^{-2}  \\
&& \times \left( \frac{\dot M}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
&&\times \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{-1/3}  \left( \frac{a}{5.2~\mbox{AU}} \right)^{-4/3} \\
&\approx & 10^{-5} \mbox{Jy} \left( \frac{d}{100\mbox{~pc}} \right)^{-2}  \\
&& \times \left( \frac{\dot M}{10^6 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \\
&&\times \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{-1/3} \left( \frac{a}{5.2~\mbox{AU}} \right)^{-4/3} %\\
%&\approx &  10^{-3} \mbox{Jy} \left( \frac{P_{\rm radio}}{200 P_{\rm radio, J}} \right) \left( \frac{l}{10\mbox{~pc}} \right)^{-2} \\
%&&\times \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{-1} \left( \frac{R_p}{R_{p, J}} \right)^{3}
\end{eqnarray}
%%%
Here, we used $P_{\rm radio,\,J}=2.1\times 10^{11}\mbox{~W}$ \citep{griebmeier2007} \memoYF{cite original paper!}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
 \begin{minipage}{0.5\hsize}
  \begin{center}
   \includegraphics[width=\hsize]{radio_emission_Mdot1e-8.eps}
  \end{center}
 \end{minipage}
 \begin{minipage}{0.5\hsize}
  \begin{center}
   \includegraphics[width=\hsize]{radio_emission_Mdot1e-5.eps}
  \end{center} 
 \end{minipage}
   \caption{Top panel: Intensity of radio emission (in the unit of Jy) from a companion to a RG star (left) and an AGB star at 1 pc. The hashed region with vertical lines are not observable because of the plasma frequency cut-off of Earth's ionosphere. The hashed region with horizontal lines are not observable because of the plasma frequency cut-off of the stellar wind plasma in the vicinity of the companion. Bottom panel: cyclotron frequency, i.e., the frequency of radio emission. }
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%\subsection{}
The peak of the auroral radio emission is around the cyclotron frequency, of the planetary magnetic field $f_{\rm cyc}$: 
%In our assumption, the frequency band width is assumed to be scaled with the cyclotron frequency, $f_{\rm cyc}$: 
%%%
\begin{eqnarray}
f_{\rm cyc} &=& \frac{eB}{2\pi m_e} \approx 28 {\rm~MHz} \left( \frac{B}{10 \rm~G} \right) \, ,
\label{eq:cyc} \\
B &=& \frac{\mu_0}{2\pi}\frac{\mathcal{M}}{R_p^3} \\
&\sim & 9.1{\rm~G} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right) \left( \frac{R_p}{R_{p, J}} \right)^{-3}. 
\end{eqnarray}
%%%
%where $\mathcal{M}_J = 1.56 \times 10^{27} \mbox{A m}^2$. 
For magnetic field strengths in the vicinity of $\sim$25-50~Gauss (G),
the cyclotron frequency will be in the range $\sim$75-150~MHz.

Note that the emission is observable from ground only when the maximum frequency is larger than the plasma frequency of the Earth's ionosphere,
%%%
\begin{equation}
f_{\rm c1}\sim 10~\mbox{MHz}
\end{equation}
%%%
{\it and} the maximum plasma frequency along the line of sight, $f_{\rm c2}$. 
At the favorable condition where the planet is in the near side of the spherical plane, the maximum plasma frequency corresponds to that in the vicinity of the planet, $f_{\rm plasma}$. Therefore, 
%%%
\begin{eqnarray}
f_{\rm c2} &=& f_{\rm plasma} = \frac{1}{2\pi} \sqrt{\frac{ne^2}{\epsilon _0 m_e}} \\
&\approx & 13{\rm~MHz} \left( \frac{a}{5~\mbox{AU}}\right)^{-1} \notag \\
&& \times \left( \frac{\dot M}{10^6 \dot M_{\odot}}\right)^{1/2} \left(\frac{v}{10^{-1} v_{\odot}}  \right)^{-1/2}
\end{eqnarray}
%%%

@@

Unlike usual hot jupiters, RGHJs are not subject to tidal lock, as the gravitational effects of their host star does not change even if the star evolves into red giants. Without no physical insights of the typical spin rate, we simply assume that of Jupiter: $\omega = 9.925$ [hr]. 
\memoYF{Would their spin rate become higher or lower??}

In order to evaluate $\rho _c $ and $r_c$, we need a model of internal structure of gaseous planets. 
First, we assume that the planetary radius is constant at $R_p = R_{p,J}$, as the numerical calculations show that the radii of gaseous planets over the range of $0.1 R_{p, J} < M_p < 10M_{p, J}$ (with core mass less than 10\%) are converged to $0.8 R_{p, J} < R_p < 1.2R_{p, J}$ in 1 Gyr \citep{fortney2007}. 
For the density profile, we assume a polytrope gas sphere with index $n=1$, which results in:
%%%%%%%%%% 
\begin{equation}
\rho [r] = \left( \frac{\pi M_p}{4 R_p^3} \right) \frac{\sin \left[ \pi \frac{r}{R_p} \right]}{\left( \pi \frac{r}{R_p} \right)}. \label{eq:rho_r}
\end{equation}
%%%%%%%%%%
We determine the core radius $r_c$ by assuming that the hydrogen becomes metallic when $\rho (r)$ exceeds the critical density $\rho_c=700\,\mbox{kg/m}^3$. The density of the metallic core, $\rho _c$ is obtained by averaging the density in the core. 
We assume that the conductivity $\sigma $ is the same as Jupiter \memoYF{?}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[htdp]
\caption{Canonical Parameters of 2 types of hot jupiters.}
\begin{center}
\begin{tabular}{c|ccc} \hline \hline
& mass & orbital distance &  spin period \\ 
& $M_p$ & $a$ &  \\ \hline
canonical HJ & $M_J$ & 0.05 AU & 4 days \\
RGHJ & $M_J$ & 5 AU &  0.4 days \\ \hline
\end{tabular}
\end{center}
\label{tab:comp_HJ}
\end{table}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

On the assumption that the radio emission is proportional to the input kinetic energy (equation (\ref{eq:Pinp_kin})), the scaling of the radio emission is expanded as follows: 
%%%
\begin{eqnarray}
P_{\rm k,\,inp} &=& nv^3 r_s^2 = n v^3 \left( \frac{\mu_0^2 \mathcal{M}^2}{8\pi^3 m_{\rm p}nv^2} \right)^{1/3}  \\
&=& P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{2/3} \left( \frac{a}{5.2~\mbox{AU}} \right)^{-4/3}  \notag \\
&& \times \left( \frac{\dot M}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \label{eq:Pkinp}
\end{eqnarray}
%%%

It is illustrative to compare radio emission power of canonical hot jupiters and that of red giant hot jupiters. We adopt canonical values for planetary mass, orbital distance, and orbital period of each type as shown in Table \ref{tab:comp_HJ}. Then, equation (\ref{eq:Pkinp}) can be re-normalized as 
%%%
\begin{eqnarray}
P_{\rm k,\,inp} 
&\approx & 230 ~P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M}{10^6 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3}  \\
&\approx & 2.3 \times 10^4 ~P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M}{10^9 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3}  \\
&\approx & 100 ~P_{\rm k,\,inp,\,J} \left( \frac{\mathcal{M}}{0.1\mathcal{M}_J} \right)^{2/3} \left( \frac{a}{0.05~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} 
\end{eqnarray}
%%%
\memoYF{what is the typical wind velocity of AGB stars?}
Thus, the jupiters to the AGB stars are intrinsically 200 times as bright as canonical hot Jupiters. In other words, we could search for radio signal from these stars in $(\sqrt{200})^3 \approx 3000$ times as large volume as the canonical hot jupiters.

We now turn to the spectral radio flux observed at the Earth. 
%%%
\begin{eqnarray}
F_{\nu} &=& \frac{P_{\rm radio}}{\Omega d^2 f_{\rm cyclotron}} \\
%&=& P_{\rm radio,\,J} \\
&\approx & 5.4\times 10^{-6} \mbox{Jy} \left( \frac{d}{10\mbox{~pc}} \right)^{-2}  \notag \\
&&\times \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{-1/3}  \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
&\approx & 1.0 \times 10^{-3} \mbox{Jy} \left( \frac{d}{10\mbox{~pc}} \right)^{-2}  \notag \\
&&\times \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{-1/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\ 
&& \times \left( \frac{\dot M}{10^6 \dot M_{\odot}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} 
%\\
%&\approx &  10^{-3} \mbox{Jy} \left( \frac{P_{\rm radio}}{200 P_{\rm radio, J}} \right) \left( \frac{l}{10\mbox{~pc}} \right)^{-2} \\
%&&\times \left( \frac{\mathcal{M}}{\mathcal{M}_J} \right)^{-1} \left( \frac{R_p}{R_{p, J}} \right)^{3}
\end{eqnarray}
%%%
Here, we used $P_{\rm radio,\,J}=2.1\times 10^{11}\mbox{~W}$ \citep{griebmeier2007} \memoYF{cite original paper!}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
 \begin{minipage}{0.5\hsize}
  \begin{center}
   \includegraphics[width=\hsize]{radio_emission_Mdot1e-8_constRp.eps}
  \end{center}
 \end{minipage}
 \begin{minipage}{0.5\hsize}
  \begin{center}
   \includegraphics[width=\hsize]{radio_emission_Mdot1e-5_constRp.eps}
  \end{center} 
 \end{minipage}
   \caption{Top panel: Intensity of radio emission (in the unit of Jy) from a companion to a RG star (left) and an AGB star at 1 pc. The hashed region with vertical lines are not observable because of the plasma frequency cut-off of Earth's ionosphere. The hashed region with horizontal lines are not observable because of the plasma frequency cut-off of the stellar wind plasma in the vicinity of the companion. Bottom panel: cyclotron frequency, i.e., the frequency of radio emission. }
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
